%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[english,twoside,a4paper,11pt]{article}
\usepackage[USenglish]{babel}
\usepackage[pdftex]{graphicx}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb}
\usepackage{url}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \setcounter{tocdepth}{0}
 \setcounter{secnumdepth}{5}
 \setcounter{totalnumber}{6}      % anzahl der floats pro seite
 \setcounter{topnumber}{6}
 \setcounter{bottomnumber}{6}
\numberwithin{equation}{section}
 \renewcommand{\topfraction}{1.0}
 \renewcommand{\bottomfraction}{1.0}
 \renewcommand{\textfraction}{0.0}
 \fboxrule0.25mm
 \unitlength 1cm
 %\voffset -0.75in
 %\hoffset -1in
%%%\headincludeon
%%%\headseplineon
 \pagestyle{headings}
 \textheight 230mm
 \textwidth 150mm
 \topmargin 0mm
 \addtolength\headsep{3mm}
 \oddsidemargin 10mm
 \evensidemargin 10mm
 \parindent 3mm
 \parskip1.5ex plus 0.5ex minus 0.3ex

\newenvironment{myindentpar}[1]%
{\begin{list}{}%
         {\setlength{\leftmargin}{#1}}%
         \item[]%
}
{\end{list}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\thispagestyle{empty}
\begin{center}
\begin{Large}
DC power flow -- internal documentation \\
\today\\
\end{Large}
{\it Rolando A Rodriguez and Sarah Becker with help from Gorm Andresen and Uffe Poulsen}
\end{center}

The following document attempts to explain the problem of the
constrained DC power flow, and the solutions implemented in the
group's code, as well as this code's algorithm. 

\section{Power flow across a fully renewable Europe}

We face the task of describing the flow of power across various nodes
in a network. These nodes represent power grids (regions or countries)
connected by high voltage transmission lines. For simplicity, we
choose to model these flows considering only Kirchhoff's laws. This
means that, based on the power generation and consumption (load) in
each node, we look for the solution to the problem how power should be
transmitted from sources to sinks which minimizes the necessary
transmission capacities. 

This approach neglects market influences, i.e. the fact that power
exchanges from country to country are directly controlled using power
electronic devices, such as FACTS or HVDC lines. We expect that such
economic agreements between countries lead to
shortest-path flows, which is not the optimal solution if one is
interested in minimizing the transmission capacities needed.

In particular, we are interested in a fully renewable scenario, where
the overall load is (on average) completely covered by generation of
wind and solar energy. For a realistic model, we use actual weather
data, which is converted into wind and solar energy generation and
then normalized to match the average load. For details about the load
and generation time series, see Dominik Heide's PhD thesis. The
important parameters in this context are $\alpha$, which is the
percentage of wind energy in the mix, and $\gamma$, which is a global
factor. Together with the load time series, we get from that a time
series of power mismatches for each country, denoted by
\begin{equation}
\begin{split}
\Delta &= \text{renewable generation} - \text{load}\\
&= \gamma\,(\alpha\cdot \text{wind power}+ (1-\alpha)\cdot\text{solar
  power}) - \text{load}.
\label{eq:powermismatch}
\end{split}
\end{equation}
From the power mismatches, we want to determine the
power flows across Europe, subject to various outer conditions like
transmission line capacities or different $\alpha$ and $\gamma$ values.

\section[From AC to DC]{The DC power flow as a simplification of the AC power flow}

Power transmission in large networks can be precisely described by AC
power flow, which accounts for voltage drops, phase differences, and
active and reactive power. Each of the $N$ buses (nodes) of the
system is described by four variables: real power mismatch $P$,
reactive power mismatch
$Q$, the voltage $V$ and the voltage phase angle $\delta$. Depending
on whether a bus is a generator bus, a load bus, or a reference
(slack) bus, different variables are used in engineering. As
capacitive lines require energizing before being able to transmit
power, and inductive loads consume reactive power, the values of $P$,
$Q$, $V$, and $\delta$ will vary throughout the network. Solving the
power flow and finding the power levels of buses and lines requires
solving two sets of equations, one for the active and one for the
reactive power.
\begin{align}
P_i&=\sum_{k=1}^N |V_i|
|V_k|(G_{ik}\cos\delta_{ik}+B_{ik}\sin\delta_{ik}) 
\label{eq:realAC} \\
Q_i&=\sum_{k=1}^N |V_i|
|V_k|(G_{ik}\sin\delta_{ik}-B_{ik}\cos\delta_{ik}) 
\label{eq:imagAC} 
\end{align}
where $\delta_{ik}=\delta_i-\delta_k$ is the difference of the phase
angles at nodes $i$ and $k$, and $G_{ik}$ and $B_{ik}$ are the
real and imaginary parts of the admittance of the links connecting
nodes $i$ and $k$, respectively. The consideration of reactive power
and voltage is integral for transmission system operators (TSOs) in
order to detect instability in transmission and distribution systems,
but complicates the solution significantly. These equations are
usually solved using the Newton-Raphson method.

For our purposes, we assume a stable system with stable voltage
levels and no active or reactive power losses. This means that the
power mismatch in each node is equal to its real part, i.e.
\begin{align*}
\Delta_i= P_i,\quad Q_i = 0
\end{align*}
In this case, we can
simplify the AC power flow to a DC power flow: Essentially, our
assumptions imply that the voltage levels throughout the grid are
constant, so that $V_i=V_k\equiv 1$, that there are no real components
in the impedance of the lines (that is, they have resistances
$R_{ik}=0$ such that $G_{ik} =0$), and the difference between the
phase angles is sufficiently small so that $\sin(\delta_{ik} )\approx
\delta_{ik}$. This means that we can forget about \eqref{eq:imagAC}, and
\eqref{eq:realAC} is simplified to
\begin{align}
\Delta_i=\sum_{k\neq i}^N B_{ik}(\delta_i-\delta_k)
=\sum_{k\neq i}^N B_{ik}\delta_i - \sum_{k\neq i}^N B_{ik}\delta_k
\equiv \sum L_{ik}\delta_k,
\label{eq:DC}
\end{align}
if we define
\begin{align*}
L_{ik} =\begin{cases} -B_{ik} & \text{ if $i\neq k$} \\
 \sum_{k\neq i}B_{ik} & \text{ if $i = k$}
\end{cases}
\end{align*}
In our case where $B_{ik}=1$, the $L$ matrix is equal to the
Laplace matrix
\begin{align*}
L=K K^T
\end{align*}
where K is the incidence matrix of the network with the elements
\begin{align*}
K_{nl} = \begin{cases}
\phantom{-} 1 & \text{ if link $l$ starts at node $n$} \\
-1 & \text{ if link $l$ ends at node $n$} \\
\phantom{-} 0 & \text{ otherwise}
\end{cases}
\end{align*}
The values of $\delta_n$ are now obtained from \eqref{eq:DC}. They
define the flows between two nodes, as the power flow $F_l$ along link
$l$ that connects nodes $i$ and $k$ is given by 
\begin{align*}
F_l =B_{ik} (\delta_i-\delta_k ),
\end{align*}
which, in the case where all the lines have $B_{ik} =1$, is reduced to
\begin{align}
F= K^T\delta
\label{eq:flow}
\end{align}
To sum up, the problem is now to solve for $\delta$ in \eqref{eq:DC}
in order to find the flows using \eqref{eq:flow} for given $\Delta_i$ and
network topology. As B does has one eigenvalue zero, this is only
possible for $\Delta$s that obey the constraint $\sum_i
\Delta_i=0$. Additionally, the solution is only unique up to a global
phase, i.e. one can fix e.g. $\delta_0=0$. Alternatively, the
Moore-Penrose pseudo inverse can be obtained for $L$, to solve
directly, the latter option saving calculation time in very large
systems. Confer Dominik Heide's dissertation for more details.

\section{The minimum dissipation principle}

We now want to include constraints on the flows that model the finite
transmission capacity of a power line, i.e. inequalities of the form
\begin{align}
F=K^T \delta \leq h
\label{eq:flowconstr}
\end{align}
Additionally, we want to treat cases in which 
\begin{align}
\sum_i \Delta_i\neq 0.
\label{eq:powerdelta}
\end{align}
This is obviously not possible in
our setting so far, since there are no solutions at all if $\sum_i
\Delta_i\neq 0$ (no surprise here -- that would violate the conservation of
energy), and if there are solutions, they are unique up to a global
phase and do not allow for further constraints on the flows. What we
need here is a reformulation of the problem that allows for the
inclusion of \eqref{eq:flowconstr} and \eqref{eq:powerdelta}. In this
section, we will restate the problem, and in the next section, we will
include \eqref{eq:flowconstr} and \eqref{eq:powerdelta}.

While we have toyed with the idea of changing the reactances of
individual lines to regulate the transmission limits, we found that
the problem is easier to solve if we state the power in the nodes
directly as a function of the flows, and forget about
$\delta$. Eq. \eqref{eq:DC} then simply becomes
\begin{align}
KF= \Delta
\label{eq:DC2}
\end{align}
We have now passed from a system in $N$ (number of nodes) variables to
a system of $L$ (number of links) variables. For a generic power
network, $L$ will larger than $N$, and therefore, the solution of this
problem is no longer unique. Additional solutions arise because
Eq. \eqref{eq:DC2} can be satisfied by flows that have, in addition to
the ``net transporting current'' that takes power from node to node,
``circular components'' that flow in a round. Fortunately, it is
possible to make it unique and identical to the DC power flow
situation by requiring that the square of the flows be minimal. This
eliminates the circular flows. It is called the minimum dissipation
principle. We show that the solution
to Eq. \eqref{eq:DC2} together with the minimization is indeed the
same as the one of Eq. \eqref{eq:DC} using a vector of $N$ Lagrange
multipliers $\lambda$. We are looking for the minimal flows under the
constraint that the power is transported from the sources to the sinks.
\begin{align}
\min_F F^T F -\lambda^T (KF-\Delta).
\label{eq:lagr}
\end{align}
It's easy to see the minimum by completing the squares in $F:$
\begin{align*}
\left(F- \frac{1}{2}K^T \lambda\right)^T 
\left(F- \frac{1}{2}K^T \lambda\right)
=F^T F-\lambda^T KF+ \frac{1}{4}\lambda^T K K^T \lambda
\end{align*}
so that Eq. \eqref{eq:lagr} becomes
\begin{align*}
\min_F
\left(F- \frac{1}{2}K^T \lambda\right)^T 
\left(F- \frac{1}{2}K^T \lambda\right)-
\frac{1}{4}\lambda^T K K^T \lambda  +\lambda^T \Delta
\end{align*}
whose minimum evidently lies at $F=1/2 K^T \lambda$, i.e.
the minimizing flows fulfill an equation of the same form as
Eq. \eqref{eq:DC} (the flow is a so-called potential flow). This means
that, as long as we minimize the square 
of the flows $F^T F$ while ensuring that $KF= \Delta$, we are describing
the DC power flow, or, in other words, the two formulations are
equivalent.

We now turn to the extension of the problem to cases where $\sum_i
\Delta \neq 0$. In order to achieve that, we introduce {\it balancing},
i.e. extra power generation when there is less renewable generation
than load, and {\it curtailment}, i.e. the shedding of
overproduction. At the same time, we include the constraints
on the flows. Computationally speaking, this is a quadratic
programming optimization problem. Numerical solutions are implemented
in a number of packages, developed for MATLAB, C and Python. % We
% describe in the following how the problem must be stated for the
% \texttt{CVXOPT} module for Python. Other implementations are quite
% similar. Along the way, we will come across the precise definitions of
% curtailment and balancing.
We describe in the following how it is done with C modules generated
by \texttt{cvxgen}.

\section{Balancing and flow calculation with \texttt{cvxgen}}

To calculate balancing needs and occurring flows, we employ the
physical DC approximation to the full AC power flow. It is valid as
long as the network is in a stable state, i.e. no significant voltage
or phase shifts between the nodes, and for a network in which line
resistances can be neglected. It has the obvious advantage that it
minimizes the total dissipation in the network, i.e. the power losses,
and the necessary flows.

We first take a look at a situation in which there is no global
mismatch between renewable energy generation and load.
\begin{align}
\sum_{i=1}^N \Delta_i =0
\label{eq:nomis}
\end{align}
For a network with unconstrained line capacities (``copper flow''),
the DC power flow is then given as that set of flows $F_l$ that
fulfills 
\begin{align}
\nonumber &KF=\Delta \qquad\qquad\quad\\
\text{ and minimizes } &\sum_{l=1}^L F_l^2.
\label{eq:minf}
\end{align}
Due to condition \eqref{eq:nomis}, the existence of such a solution is
guaranteed, and the minimization \eqref{eq:minf} makes it unique. 

Now we generalize this to the case where we have a global mismatch,
which has to be compensated by balancing. The total balancing can be
expressed as\footnote{$(x)_-=\max\{-x,0\}$ denotes the
  negative part of a quantity $x$.}
\begin{align*}
B_{\rm tot}=\sum_{i=1}^N(\Delta_i-(K\cdot F)_i)_-=\sum_{i=1}^N B_i
\end{align*}
The balancing need $B_i$ is what is potentially left of a negative
mismatch $\Delta_i$ after it has been reduced by the net imports
$-(K\cdot F)_i$. We use the flows to minimize $B_{\rm tot}$. In other
words, we make the system use as much of the 
renewable generation as possible. This is our top priority.

As a secondary objective, we still want to minimize dissipation. We do
so in a second step, where we minimize the flows keeping the total
balancing at its minimal value found in the first step:
\begin{align*}
\min_{\sum_{i=1}^N (\Delta_i-(K\cdot F)_i)_- \leq B_{\rm min}}\sum_{l=1}^L F_l^2
\end{align*}
This ensures that we arrive at the minimal flows that make optimal
sharing of renewables possible.

If the line capacities are constrained by (possibly direction
dependent) capacities, $h_{-l}\leq F_l\leq h_l$, these constraints are
included in the two minimizations, and we have
\begin{align*}
\text{Step 1: }&\qquad \:\:
\min_{h_{-l}\leq F_l \leq h_l} \qquad \:\: 
\sum_{i=1}^N(\Delta_i-(K\cdot F)_i)_- = B_{\rm min}\\
\text{Step 2: }&
\min_{\substack{h_{-l}\leq F_l \leq h_l \\ \sum_{i=1}^N (\Delta_i-(K\cdot F)_i)_-
  \leq B_{\rm min}}} \sum_{l=1}^L F_l^2
\end{align*}

\section{Balancing, flow and storage calculation with \texttt{gurobi}}

...comes as soon as I understand it.

\section{Other tested implementations}

\texttt{CPLEX} -- commercial, apparently a predecessor of gurobi,
slower

\texttt{cvxopt} -- python-internal module, easy, but very slow. In
addition, defining a hierarchy of goal just by giving the higher
priority cost function a higher prefactor wasn't such a good idea --
in the end, the routine just ignored the secondary targets because
their influence on the cost function was too low.

% \section{Implementing balancing and curtailment}
% \label{sec:3}

% The quadratic programming solver \texttt{CVXOPT} (from ConVeX
% OPTimisation) requires the information to be stored in a specifically
% shaped matrices. The form and values of these matrices help define the
% scenario and problem that one wants to optimize. The standard format of
% a quadratic programming problem is
% \begin{align}
% C&=x P x+q x \text{ -- the cost function to be minimized}
% \label{eq:cost} \\
% A x&=b \text{ -- the equality constraint}
% \label{eq:eq}\\
% G x &\leq h \text{ -- the inequality constraint}
% \label{eq:ieq}
% \end{align}
% where $x$ is the variable that is varied and the other entities are
% parameters of the specific problem (don't confuse the cost matrix $P$
% with the power we talked about earlier!). In our case,
% $x$ does not only consist of the flows $F_l$, but we also want to
% minimize balancing $B_n$ and curtailment $R_n$ at each node. This
% means that our $x$ looks like
% \begin{align*}
% x=\left(\begin{matrix}
% F_1\\
% \vdots\\
% F_L\\
% B_1\\
% \vdots\\
% B_N\\
% R_1\\
% \vdots\\
% R_N
% \end{matrix}
% \right)
% \end{align*}

% Let's first take a look at the cost function. The quadratic parameter
% $P$ is a $(L+2N)\times(L+2N)$ matrix. It is simply chosen to be
% diagonal, since there is no distinguished link or node in the network
% and we do not want to introduce any ``synergy effects'' between flows,
% balancing and curtailment. We thus have:
% \begin{align*}
% P=\left(
% \begin{matrix}
% P_F & 0 & 0 \\
% 0 & P_B & 0 \\
% 0 & 0 & P_R
% \end{matrix}
% \right)
% \end{align*}
% Upon examination, we found that in Dominik Heide's code, curtailment is
% often shared between nodes. That is, in the event of a global positive
% mismatch where node $i$ has an overproduction of $R$ and node $j$ is
% perfectly balanced, Dominik's algorithm will have node $i$ export some
% of its overproduction to node $j$, and both will curtail amounts
% $0.5\cdot R$. This behavior is of course not desired since it leads to
% additional flows which are completely superfluous. The reason for
% these results is that Dominik minimized the square of the balancing
% and treated curtailment just as negative balancing. If you do this,
% however, an even distribution of curtailment is favored, leading to
% transport of power which is eventually shedded. This is why we treat
% curtailment and balancing separately, and this is also the reason why
% the quadratic parameter $P_R$ is set to zero.

% The cost of curtailment is instead accounted for in the linear
% term. Again, since no nodes or links are distinguished, it is chosen
% to look like
% \begin{align*}
% q=\left(
% \begin{matrix}
% q_F \\
% q_B \\
% q_R
% \end{matrix}
% \right)
% \end{align*}

% Next, we build the equality constraint matrix $A$. It should
% incorporate the DC power flow equation Eq. \eqref{eq:DC2}, which must
% now be modified to include balancing and curtailment. It should now
% express
% \begin{align*}
% \text{renewable generation}+
% \text{balancing}=\text{load}+\text{curtailment} + \text{net outflow}.
% \end{align*}
% Since load and generation are, in fact, our input parameters, they
% form the outer constraint on our system, and we have
% \begin{align*}
% (\text{renewable generation}-\text{load})_n=\Delta_n
% =\sum_l F_{ln}-B_n+R_n.
% \end{align*}
% This means that $A$ is given by
% \begin{align*}
% A=\left(
% \begin{matrix}
% K & -\mathbb{I}_{N\times N} & \mathbb{I}_{N\times N}.
% \end{matrix}
% \right)
% \end{align*}
% and $b=\Delta$ by the nodal mismatch of generation and load.

% Finally, the inequality constraints must follow the form $G x\leq
% h$. Remembering that flows are directed and that the link capacity
% can have different values in either direction, the flow component of
% the constraints must have the form
% \begin{align}
% \left(
% \begin{matrix}
% 1 & 0 & \hdots \\
% -1 & 0 & \hdots \\
% 0 & 1  & \hdots \\
% 0 & -1  & \hdots \\
% \vdots & \vdots & \ddots 
% \end{matrix}
% \right)
% \left(
% \begin{matrix}
% F_1 \\
% F_2 \\
% \vdots \\
% \vdots
% \end{matrix}
% \right)
% \leq
% \left(
% \begin{matrix}
% h_1 \\
% h_{-1} \\
% h_2 \\
% h_{-2}\\
% \vdots
% \end{matrix}
% \right),
% \label{eq:capc}
% \end{align}
% where $h_l$ is the capacity flowing in the positive direction along
% link $l$ and $h_{-l}$ is the capacity flowing in the opposite
% direction, with $h_l, h_{-l} \geq 0$.

% Both the curtailment and the balancing are expressed in positive
% values, so that both have to limited with the condition that $B_n, R_n
% \geq 0$. This implies that the constraint G matrix in the
% balancing and section is defined as
% \begin{align}
% -\mathbb{I} B \leq 0
% \label{eq:bal+}
% \end{align}

% For the curtailment, we have the additional constraint that there
% should be no transported curtailment at a node, i.e. it must be
% limited to the local excess power. Together with the positivity, this
% means for the curtailment part of $G$:
% \begin{align}
% \left(
% \begin{matrix}
% 1 & 0 & \hdots \\
% -1 & 0 & \hdots \\
% 0 & 1  & \hdots \\
% 0 & -1  & \hdots \\
% \vdots & \vdots & \ddots 
% \end{matrix}
% \right)
% \left(
% \begin{matrix}
% R_1 \\
% R_2 \\
% \vdots \\
% \vdots
% \end{matrix}
% \right)
% \leq
% \left(
% \begin{matrix}
% b_1 \\
% 0 \\
% b_2 \\
% 0\\
% \vdots
% \end{matrix}
% \right)
% \label{eq:cur+}
% \end{align}

% \section{Different balancing strategies}
% \label{sec:4}

% Given that our solution vector $x$ contains information about not only
% the flows, but also about the balancing and curtailment activities of
% each node, the optimization is not only minimizing the square of the
% flows, but also the mismatch correction at each node. The standard QP
% cost function
% \begin{align*}
% C=xPx+qx
% \end{align*}
% reads in detail
% \begin{align*}
% C=\sum_l \left[P_F F_l^2+ q_F F_l\right]
% +\sum_n \left[P_B B_n^2+ q_B B_n + P_R R_n^2 +q_R R_n\right]
% \end{align*}
% $P_F$ , $P_B$ and $P_R$ are the flow, balancing and curtailment
% prizing in the quadratic component of $C$ and $q_F$ , $q_B$ and $q_R$
% are linear ones. As we are interested in minimizing the square of the
% flows and not in minimizing the flow along any one link in
% particular, the linear cost of $F$ is set to zero ($q_F=0$). Likewise,
% since we don't want to minimize the square of the
% curtailment (as explained above), we set $P_R$ to zero as well. This
% leaves us with
% \begin{align*}
% C=\sum_l P_F F_l^2 +\sum_n \left[P_B B_n^2 +q_B B_n+q_R R_n \right]
% \end{align*}
% By adjusting the remaining parameters, we can motivate one kind of
% behavior over another. If, for instance, local balancing is cheaper
% than flows, countries will prefer to balance their own loads, even if
% their neighbors have excess power. Here, the parameters take different
% values, depending on which scenario we wish to model.

% \subsection{Weakly independent nodes}

% This models the case where the nodes have to do their own balancing,
% and receive no help from their neighbors apart from receiving their
% renewables. The rules to follow are (in descending priority):
% \begin{enumerate}
% \item Sharing of renewable surplus
% \item Own balancing
% \item Own curtailment
% \end{enumerate}
% To remove the motivation to share balancing, $P_B$ is set to zero. In
% order ensure that the renewable surplus is shared, the marginal cost
% of transmitting one unit of energy between any two nodes $i$ and $j$
% has to be lower than the marginal cost of producing it at the
% destination with balancing plus the cost of curtailing it at the
% origin. In other words 
% \begin{align}
% \frac{\partial C}{\partial F_{ij}} \stackrel{!}{<}
% \frac{\partial C}{\partial B_j} +\frac{\partial C}{\partial R_i}
% \label{eq:trans1}
% \end{align}
% which is in every case fulfilled if
% \begin{align*}
% 2P_F ( N-1) \Delta_0 < q_B + q_R,
% \end{align*}
% where the term $(N-1)\Delta_0$ represents the largest possible flow in
% the network, built from the largest amount of mismatch $\Delta_0$
% transferred at a given time and the maximum number of links that this flow
% could travel through. In the quadratic parameters, we usually set $P_F
% =1$. Theoretically, it shouldn't matter how $q_B$ and $q_R$ are chosen
% in order to satisfy the inequality, as any balancing will automatically
% imply curtailment if transfer could have been an option. To play
% safe, we set both of them equal to the left side of the inequality.

% \subsection{Weakly dependent nodes}

% This models the case where nodes share the balancing duties whenever
% there is a global negative mismatch, as opposed to leaving the
% countries with a negative mismatch to balance it alone. While the total
% balancing energy used will be the same in both cases, the total
% balancing power installed in Europe will be lower in this case (though
% perhaps higher for an individual, stable node than it would be in an
% uncooperative scenario). The set of
% rules is then
% \begin{enumerate}
% \item Sharing of renewable surplus
% \item Sharing of balancing
% \item Own balancing
% \item Own curtailment
% \end{enumerate}
% In this scenario, we wish to reduce the
% square of the balancing required in total, so that two countries
% balancing a mismatch $M/2$ is preferable to a single one balancing
% $M$. This implies bringing back the quadratic term $P_B$. The
% condition that flow must be cheaper than balancing still holds, so the
% inequality in \eqref{eq:trans1} still applies to this case. In
% addition, we want to arrange it so that marginal cost of balancing at a
% node $i$ which is already doing a large amount of balancing is higher
% than the marginal cost of producing and bringing balancing from a less
% used node $j$. That is
% \begin{align*}
% \frac{\partial C}{\partial B_i} 
% &> \frac{\partial C}{\partial B_j}+\frac{\partial C}{\partial F_{ij}}\\
% \text{where } \frac{\partial C}{\partial B_i}&=2P_B B_i +q_B \\
% \text{and } \frac{\partial C}{\partial B_j}+\frac{\partial C}{\partial F_{ij}}
% &=2P_B B_j +q_B+\sum_l  2P_F\Delta_0 \leq 2P_BB_j+q_B +2P_F(N-1)\Delta_0
% \end{align*}
% where, again, $(N-1)\Delta_0$ represents the maximal transferable
% mismatch along the longest possible route. This inequality can
% be simplified to
% \begin{align}
% \nonumber 2P_B B_i+q_B &> 2P_B B_j+q_B +2P_F( N-1)\Delta_0\\
% \Leftrightarrow
% P_B & > \frac{( N-1) P_F\Delta_0}{ B_i - B_j}
% \label{eq:bals}
% \end{align}
% This means that, the more uneven the current balancing productions
% between nodes $i$ and $j$ are, the (relatively) cheaper it will be to
% transport power from $j$ to $i$ in comparison to balancing at
% $i$. This term explodes as the balancing production in the two nodes
% equals out, and $B_i=B_j$, so some tolerance level must be
% introduced. We set this to $\Delta B=B_i - B_j\approx 1\rm\,MW$, and
% then set $P_B$ accordingly.

% We found that ensuring that inequality \eqref{eq:trans1} holds makes
% the system behave as a weakly independent grid, and by further
% asserting inequality \eqref{eq:bals}, we can make the system behave as
% a weakly dependent grid.

\section{Code walk-through and instructions}

The following is a walk-through of the zdcpf.py script, as far
as the version of \today \ is concerned. The code works with data
stored on the group's Pepsi server, which requires a user name
and password. Access is currently only for group members. Contact us for more information, and maybe we can come to an agreement. In order to access
the database from outside the institute, just open an ssh tunnel to
the database by typing 

\noindent \texttt{ssh -L5432:localhost:5432 USERNAME@pepsi.imf.au.dk}

\noindent into your terminal. Note that due to a brute force attack on
the server, it was necessary to switch off the password identification
when you login from a computer outside the Aarhus mathematics institute
(IMF). You have to generate an ssh key pair, e.g. by running

\noindent \texttt{ssh-keygen -t rsa -b 4096}

and put the public key on pepsi with

\noindent \texttt{scp ~/.ssh/id\_rsa.pub
  <username>@pepsi.imf.au.dk:.ssh/authorized\_keys}

\noindent (obviously, you have to do so from a computer which can
still reach pepsi by password). Note that this overwrites any existing
authorized keys. You can optionally specify a passphrase for your
private key. Your (local) private key \texttt{~/.ssh/id\_rsa} is then required
to log in, so copy it to any other computers you need to ssh to pepsi
from.

The code is written in
Python and requires a number of modules to be installed for proper
execution. These are: \texttt{numpy}, \texttt{sqlalchemy},
\texttt{psycopg2}, \texttt{cvxopt}, \texttt{pylab},
\texttt{matplotlib}, and \texttt{scipy}. Additionally, the group has
been working under \texttt{ipython}, which somewhat simplifies the
process of running scripts. This makes tests easier, but slows python
down a bit and there is a known memory leak when generating a lot of
histograms, what we actually do (see
\url{http://old.nabble.com/memory-usage-%28leakage-%29-in-ipython-interactive-mode}
\url{-td22345056.html}). So don't use it generally.

First, you need to fetch the data you need from the server. You will
find the necessary functions for this in the file
\texttt{Database\_v2.py}. There you find for example the function 
\texttt{get\_ISET\_country\_data}, which can be used to load data
(wind and solar energy generation, load etc.) of a specific
country. It first attempts to access the data locally by looking for a
\texttt{data} folder, and if the data are not found, it tries to
download them from the Pepsi server. These files are in numpy's own
binary format, and contain the 70128 data points of load time series,
as well as normalized values for offshore and onshore wind and solar
generation for each region.

Once you have the data, you will want to determine the flow time series
in a specific setting. The bulk of the code for this purpose can be
found in \texttt{zdcpf.py}. To execute, it needs additional input data
in a folder which is by default called \texttt{settings}:
\texttt{admat.txt}, in which the line capacities 
between the European countries are listed as a matrix (data from
ENTSO-E), and \texttt{ISET2ISO\_country\_codes.npy}, which is a binary
file that contains the mapping of the ISET and ISO country codes to
each other.

Before running the code, we take a look at \texttt{zdcpf.py}. In the
core of the algorithm stands the node class, defined in the first
lines of the file. This is a type of object which we define
to contain several vectors, matrices, and methods. When you declare a
variable to be of type node, you must provide it with a path to a
file, the file name, and an ID number. For example, defining a node
\texttt{Norway} using data from N.npy would be done simply by saying

\noindent
\texttt{Norway=node('./data','N.npy',0)}. 

The \texttt{Norway} variable now holds
information on Norway's load, wind and solar production and values for
gamma and alpha. Though it is technically possible to modify or
extract the information directly from this object (for example, by
typing \texttt{print Norway.wind} or \texttt{Norway.alpha=0.5}), doing so will
bring up problems since when some variables like e.g. $\alpha$ are
changed, other variables like e.g. the total renewable generation has
to be changed as well to keep the 
data consistent. Instead, it is better to use the
functions that we have defined for these purposes, and which can be
looked up in the class definition (for example \texttt{print
  Norway.getwind()} and \texttt{Norway.setalpha(0.5)}).

For convenience, there is also a class \texttt{Nodes}, which holds the
set of all nodes in a specific calculation. Its default behavior is to
include all European nodes upon construction.

Next in the script, we meet the function \texttt{AtoKh}, which
generates the incidence matrix $K$ and the flow capacity vector $h$
(as of today) from the data in \texttt{admat.txt}. The node setup may
be changed by changing the \texttt{admat.txt} file. Note, however,
that you still must provide a data file for the each node and add it
to the list of nodes in the script, or else the code will fail.

\noindent
\texttt{F = sdcpf(Nodes,admat,path\_to\_admat,copper,lapse,b,h)}

executes the actual calculations for each point in the time
series. The function receives Nodes, which is an instance of the
\texttt{Nodes} class, the name of the admat file as well as its
location, \texttt{copper} which is a
switch to turn Europe into a copper plate, i.e. consider the
transmission capacities as unlimited, \texttt{lapse} which is the
number of time steps to consider, \texttt{b}, the scaling factor for
the line capacities (default is one), and a line capacity constraint
vector $h$ (defaults to today's capacities). It modifies the Nodes to
contain the balancing and curtailment as calculated and returns the
list of flows \texttt{F}.
With these building blocks, one can simulate a series of scenarios,
create loops that cycle through different $\alpha$ or $\gamma$ values,
or constraints, or all.

While this guide is not an user's guide, and does not aim at giving a
full explanation of every line, we outline a simple example. One would,
after opening an ssh tunnel to Pepsi, open ipython and cd to the
directory where all the files are stored, then type

\noindent \texttt{\%run Database\_v2.py \\
get\_ISET\_country\_data('DK')
}

This will check whether the data file for Denmark is available and, if
it is not, download the data files for all European countries.

Then one can load all the functions for the power flow by 

\noindent \texttt{\%run zdcpf.py
}

\noindent and then type

\noindent \texttt{N=Nodes()\\
F=sdcpf(N)
}

\noindent to start a default run. Alternatively, one can set some of the
arguments of zdcpf to other values to modify the run to what one is
interested is. For example, to sweep over various $\gamma$ values, one
could type 

\noindent \texttt{N=Nodes()\\
gammas=[0.25,0.5,0.75]\\
for i in range(len(gammas)):\\
}
\indent
\texttt{N.set\_gammas(gammas[i])}\\
\indent
\texttt{F = sdcpf(N)}

Note that each iteration overwrites the previous 70128 points in the
Nodes and F variables, and so they must be saved at every
iteration. For doubts, questions, comments, complaints and threats,
email us @ {\verb rar@imf.au.dk }, or {\verb becker@fias.uni-frankfurt.de }

\end{document}
